Single-Step Hybrid Prediction Report

### 2016-09-20 reports/agronomic_data.Rmd
Update references in report

# Agronomic Data
```{r echo = FALSE, message = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "gridExtra", "modelr", "caret",
               "knitr", "tibble")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = FALSE, 
                      warning = FALSE, echo = FALSE)
```

>>>>

# Agronomic Data
```{r echo = FALSE, message = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "knitr", "tibble")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, 
                      warning = FALSE, echo = FALSE)
```

### 2016-09-20 reports/data_overview.Rmd
Update references in report

# Data Overview
```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "knitr", "tibble", "VennDiagram", "grid",
               "gridExtra", "purrr", "htmlTable")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = FALSE, 
                      warning = FALSE, echo = FALSE)
```

```{r Load_Agro_Data, message = FALSE, echo = FALSE}
pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
genos <- readRDS("./data/processed/common_genotypes.RDS")

full_hybrid <- genos$Hybrid
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% full_hybrid) %>%
  select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))
traits <- all_pheno %>%
  .$Trait %>%
  levels()

@@

```{r Geno-Number, echo = FALSE}
geno_tab_cap <- paste("Number of genotypes for Dent and Flint parent lines and",
                      "their hybrid progeny for which genomic and/or",
                      "transcriptomic data are available.",
                      "All genotypes for which transcriptomic data are", 
                      "available, also have genomic data.",
                      "Agronomic data are available for all 1,521 hybrid",
                      "genotypes.")
# Number of genotypes for either SNPs or mRNAs separated by hybrids and their
# parental Dent and Flint lines, respectively.
genos %>% 
  at_depth(.depth = 2, .f = length) %>%
  bind_rows() %>%
  select(-agro) %>%
  mutate(Type = names(genos)) %>%
  select(Type, snp, mrna) %>%
  rename("Genomic Data" = snp,
         "Transcriptomic Data" = mrna) %>%
  knitr::kable(caption = geno_tab_cap, booktabs = TRUE)
```

>>>>

# Data Overview
```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "knitr", "tibble", "purrr")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

```{r Load-Agro-Data, message = FALSE, cache = TRUE}
pheno <- readRDS("./data/processed/Pheno_stage2.RDS")
genos <- readRDS("./data/processed/common_genotypes.RDS")

full_hybrid <- genos$Hybrid
all_pheno <- pheno %>%
  rename(Observed = EST) %>%
  as_data_frame() %>%
  filter(G %in% full_hybrid) %>%
  dplyr::select(G, Observed, Trait) %>%
  filter(Trait != "ADL") %>%
  mutate(Trait = fct_recode(Trait, 
    "FAT" = "FETT",
    "DMY" = "GTM",
    "DMC" = "GTS",
    "PRO" = "RPR",
    "SUG" = "XZ"),
         Trait = fct_relevel(Trait, 
    c("DMY", "DMC", "ADF", "FAT", "PRO", "STA", "SUG")
    ))
traits <- all_pheno %>%
  .$Trait %>%
  levels()

@@

```{r Geno-Number, echo = FALSE}
geno_tab_cap <- paste("Number of genotypes for Dent and Flint parent lines and",
                      "their hybrid progeny for which genomic and/or",
                      "transcriptomic data are available.",
                      "All genotypes for which transcriptomic data are", 
                      "available, also have genomic data.",
                      "Agronomic data are available for all 1,521 hybrid",
                      "genotypes.")
# Number of genotypes for either SNPs or mRNAs separated by hybrids and their
# parental Dent and Flint lines, respectively.
genos %>% 
  at_depth(.depth = 2, .f = length) %>%
  bind_rows() %>%
  dplyr::select(-agro) %>%
  mutate(Type = names(genos)) %>%
  dplyr::select(Type, snp, mrna) %>%
  rename("Genomic Data" = snp,
         "Transcriptomic Data" = mrna) %>%
  knitr::kable(caption = geno_tab_cap, booktabs = TRUE)
```

### 2016-09-20 reports/index.Rmd
Update references in report

---
title: Single-step Hybrid Predition
author: Matthias Westhues, Claas Heuer
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
output:
  bookdown::gitbook: default
documentclass: book
bibliography: ["library.bib", "cheuer.bib"]
biblio-style: apalike
link-citations: yes
---

# Preface {-}
This book covers an extension of the methodology of single-step prediction 
&mdash; which was developed for the imputation of SNP data from pedigree 
information &mdash; to the imputation of quantitative predictor variables using
genomic information.

@@

>>>>

---
title: Single-step Hybrid Predition
author: Matthias Westhues, Claas Heuer
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
output:
  bookdown::gitbook:
documentclass: book
bibliography: ["cheuer.bib"]
biblio-style: apalike
link-citations: yes
---

# Preface {-}
This report covers an extension of the methodology of single-step prediction 
&mdash; which was developed for the imputation of SNP data from pedigree 
information &mdash; to the imputation of quantitative predictor variables using
genomic information.

Here, we are working with a "full" data set &mdash; referring to the maximum 
number of hybrids for which agronomic and genomic, but not necessarily 
transcriptomic data are available &mdash; and a "reduced" data set with less 
hybrids but all of their parental inbred lines having complete genomic and 
transcriptomic records.

All our predictive abilities were evaluated using a leave-one-out cross 
validation (LOOCV) routine and results are displayed for T0-hybrids only.
Note, that T0-hybrids are defined as hybrids having zero parents that have been
previously tested in other hybrid combinations.

@@

## Objectives
1.    Assess whether transcriptomic data with missing records for some genotypes         can be imputed successfully based on genomic information to improve 
      prediction of unseen hybrids.
2.    Explore the utility of various machine learning algorithms for the 
      prediction of unseen hybrids.^[This is motivated by the fact that the 
      success of hybrid prediction is a strongly convex function of predictive 
      ability.]

## Overview
Chapter \@ref(introduction) gives a summary of the relevant literature.

Chapter \@ref(data-overview) provides a brief summary of the available data 
without delving into their properties.

Chapter \@ref(agronomic-data) gives an overview of the distribution of the 
agronomic traits as well as correlations among them, separately for the full
set of hybrids as well as the reduced set of hybrids.

Chapter \@ref(prediction-for-reduced-data) deals with the results of hybrid 
prediction for the reduced set of data.

Chapter \@ref(prediction-with-imputed-mrnas) examines the influence of certain levels of missingness 
in the transcriptomic data.

Chapter \@ref(conclusions-and-outlook) provides a summary of the analyses and 
provides an outlook for future analyses.

### 2016-09-20 reports/loocv_reduced.Rmd
Update references in report

# Prediction For Reduced Data
```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "gridExtra", "modelr", "MASS")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE,
                     verbose = FALSE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6, echo = FALSE, 
                      warning = FALSE, echo = FALSE)
```

@@

## Predictive ability for non-imputed data
Peak at predictive abilities based on non-imputed predictor information and rank
the results by their predictive ability.
```{r Non-Imputed-Pred-Ability, cache = TRUE}
loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  mutate(Predictor = as.factor(Predictor)) %>%
  droplevels() %>%
  group_by(Predictor, Trait) %>%
  summarize(Pred_Ability = cor(y, yhat)) %>%
  arrange(desc(Pred_Ability)) %>%
  rename("Predictive Ability" = Pred_Ability) %>%
  knitr::kable()
```

>>>>

# Prediction For Reduced Data
```{r message = FALSE, echo = FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("data.table", "ggplot2", "ggthemes", "dplyr", "forcats", 
               "viridis", "tidyr", "modelr")

knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",", digits = 2)
})
knitr::opts_knit$set(root.dir = "~/gamazon/",
                     inline = TRUE,
                     verbose = FALSE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6,
                      warning = FALSE, echo = FALSE)
```

@@

## Predictive ability for non-imputed data
Peak at predictive abilities based on non-imputed predictor information and rank
the results by their predictive ability.
```{r Non-Imputed-Pred-Ability, cache = TRUE}
loocv_tbl_cap <- paste("Predictive abilities for the 'reduced' set of hybrids",
                       "using two different predictors and seven agronomic",
                       "traits. Results are ranked by predictive ability.")
loocv_clean %>%
  filter(Dent_NA_Fraction == 0, Flint_NA_Fraction == 0) %>%
  mutate(Predictor = as.factor(Predictor)) %>%
  droplevels() %>%
  group_by(Predictor, Trait) %>%
  summarize(Pred_Ability = cor(y, yhat)) %>%
  arrange(desc(Pred_Ability)) %>%
  rename("Predictive Ability" = Pred_Ability) %>%
  knitr::kable(caption = loocv_tbl_cap, booktabs = TRUE)
```
